Stack Safety is a Security Property

### 2021-04-15 tex/main.tex
redid pass-by-ref example

\begin{figure}
  \centering
  \begin{subfigure}{.45\textwidth}
\begin{verbatim}
int *stash = 0;

int main() {
0:         int x, y;
1[C(x)]:  f(x);
2:         x = 0;
3[C(y)]:  f(y);
4[R]:      return x;
}

void f(int &a) {
10:    if(!stash)
11:      stash = a;
12:    *stash = 5;
13[R]: return;
}
\end{verbatim}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l}
      {\tt 10} &
      \memoryaddrs[(0)]{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 13} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}
      \memory{1}{\mainsealc}
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{5}
      \\
      {\tt 3} &
      \memoryaddrs[(0)]{12em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 10} &
      \memoryaddrs[(0)]{17em}
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 13} &
      \memoryaddrs[(0)]{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\retptrc}
      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{\bf 5}
\end{tabular}
\end{subfigure}
\caption{A violation with pass-by-reference}
\label{fig:passing}
\end{figure}

@@

The same annotation mechanism can be used to pass
variables by reference, and here there are important stack safety concerns,
as seen in Figure~\ref{fig:passing}. In this example, where we revert to assuming
that arguments are passed in registers,
{\tt main} calls {\tt f} with a reference to {\tt x}, but then clears {\tt x} and instead
passes a reference to {\tt y}. We would expect that the first call to {\tt f} might modify
{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
{\tt x} in the second call.

So at the initial call, {\tt x} is passed, and {\tt f} is free to write to it. Then upon
return, {\tt main}'s frame is unsealed. At the second call it is {\tt y} that is passed,
and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
a future call with no arguments passed by reference would also violate integrity if it wrote
to either variable.

When passing by reference, it is possible to repeatedly pass the same reference further
and further down the stack. Note, however, that this sharing of the reference must be over
an unbroken chain of calls. So it suffices to track two depths for each passed
cell: the original owner, and the highest depth it has been passed to. It is considered
accessible anywhere between the two.

>>>>

\begin{figure}
  \centering
  \begin{subfigure}{.2\textwidth}
{\small
\begin{verbatim}
int *stash = 0;

int main() {
  int x, y;
  f(x);
  x = 0;
  f(y);
  return x;
}

void f(int &a) {
  if(!stash)
    stash = &a; 
  *stash = 5;
  return;
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}{.49\textwidth}
{\small
\begin{verbatim}
   stash: .word 0
1  main:  add $2,%sp      ; allocate frame
2         mov %sp,%ra
3         add -2,%ra      ; ra = &x
4         call f          ; SHARE: %sp-2
5         mov $0,-2(%sp)  ; x = 0
6         mov %sp,%ra
7         add -1,%ra      ; ra = &y
8         call f          ; SHARE: %sp-1
9         mov -2(%sp),%ra ; set return value
10        ret

20 f:     cmp stash, $0   ; compare
21        breq #13        ; branch if equal
22        mov %ra,stash   ; stash = &a
23        jmp #15
24        mov $5, (stash) ; *stash = 5
25        ret 
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}{.29\textwidth}
    \begin{tabular}{l l}
      {\tt 20} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0,1)}}]%
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$ \\
      {\tt 25} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainpassc}
      \memory{1}{\mainsealc}
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{5}
      \\
      {\tt 8} &
      \memoryaddrs[(0)]{12em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 20} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainsealc}[{\makebox[0pt]{Seal(0)}}]%
      \memory{1}{\mainpassc}[{\makebox[0pt]{P(0-1)}}]%
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{0}
      \\
      {\tt 25} &
      \memoryaddrs[(1)]{17em}
      \memory{1}{\mainsealc}
      \memory{1}{\mainpassc}
      \memory{1}{\retptrc}
%      \memory{1}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-18em}{0.75em}{\bf 5}
\end{tabular}
\end{subfigure}
\caption{A violation with pass-by-reference\apt{SP(n) depths looked wrong; SNA double-check that I fixed them right.}}
\label{fig:passing}
\end{figure}

@@

The same annotation mechanism can be used to pass
variables by reference, and here there are important stack safety concerns,
as seen in Figure~\ref{fig:passing}. In this example, written in C++ notation,
{\tt f} takes a single argument by reference.
{\tt main} calls {\tt f} passing a reference to {\tt x}, then clears {\tt x} and calls
{\tt f} again passing a reference to {\tt y}.
We would expect that the first call to {\tt f} might modify
{\tt x}, but not the second. Unfortunately {\tt f} has hidden away our reference and modifies
{\tt x} in the second call. (Incidentally, this is perfectly compilable C++!)

To avoid cluttering the picture at the assembly level, we revert to assuming that the
argument {\tt a} is passed in a register, namely {\tt \%ra}.
The question is how the local variables {\tt x} and {\tt y} themselves
should be labelled in order for the policy to detect an integrity violation.
At the initial call, the address of {\tt x} is marked as passed, and {\tt f} is free to write to it. Then upon
return, {\tt main}'s frame is unsealed. At the second call it is the address of {\tt y} that is marked as passed,
and when {\tt f} instead tries to write to {\tt x}, it violates integrity. For that matter
a future call with no arguments passed by reference would also violate integrity if it wrote
to either variable.

When passing by reference, it is possible to repeatedly pass the same reference further
and further down the stack. Note, however, that this sharing of the reference must be over
an unbroken chain of calls. So it suffices to track two depths for each passed
cell: the original owner, and the highest depth it has been passed to. The policy considers
it to be accessible anywhere between the two.
